%%
% germanys turn.tex
%
% @author: Robin Suter
%%

\section{Germany's Turn}

\subsection*{Why did Einstein write a letter to the American president in 1939 and what was the
reply?}
In \emph{1933}, Einstein went to America to the \emph{Institute for Advanced Studies} in Princeton.
Following the discoveries of Meitner's, he wrote a letter to the White House, in which he informs about the possibilities of uranium as an energy source. He also warned the White House that this could be used to build powerful bombs.

The reply was a brush-off, they ignored his "warning". 

\subsection*{Why did the USA not pursue plans to build a bomb, while the Germans did?}
While in \emph{1944} the British authorities were convinced that a practical bomb could be built, they sent a message to the USA, but \emph{Lyman J. Briggs}, who was a conservative man and responsible for the \emph{U.S. atomic bomb developement}, ignored that letter and kept it secret.

Germany, in the midst of \emph{World War II}, was unterstandably much more excited about a possible weapon of mass destruction (who can blame them? They had a rough time...)

\subsection*{Who was Heisenberg and why did he get in trouble with the SS?}
\emph{Werner Heisenberg} was the world's greatest physicist after Einstein (famous for his work in quantum mechanics).

In \emph{1927}, another physicist (Johannes Stark), was a jerk and told the SS that Heisenberg wasn't patriotic enough, worked with the Jews, etc. etc.. (He wasn't completely wrong, Heisenberg did work with Bohr and Einstein, who were Jews, but he was loyal to Germany)

\subsection*{What saved him?}
Heisenberg's mother knew \emph{Himmler's} mother, so she got a letter from Heisenberg to Himmler. Through this family connection, the investigation from the SS ended (on the condition that he disavowed Einstein and other Jewish scientists).

In \emph{1940} Heisenberg wrote a report on how to construct a workable atomic bomb

\subsection*{What was Heisenberg’s bomb design based on?}
His bomb design was based on the results from Meitner and \emph{Fermi's} idea, that slowing down neutrons would increase a nuclear reaction. With "heavy water", water mixed with deuterium, these neutrons could be slowed down.

\subsection*{What happened at Heisenberg’s first test in early 1941?}
Nothing happened. There wasn't enough uranium to get a reaction going.

\subsection*{How did the Germans produce Uranium dust?}
Machining uranium into uranium dust required a lot of labor and is dangerous to workers. But hey, Germany had all these people in concentration camps, why not use them? Of course they did, they used female "slaves" to prepare the uranium oxide. Yaay.

\subsection*{When did the Germans get the first successful test results?}
In spring \emph{1942}, the breakthrough happened. It was pouring out 13 percent more neutrons than the source had pumped in. This made Himmler happy.

\subsection*{When did the Americans start to develop their bomb and what was the name of the
project?}
Einstein heard of the work in Germany and wrote another letter to the President, \emph{Franklin D. Roosevelt}. But Einstein was a socialist, so the americans didn't like him much. They ignored everything he said.

In \emph{1941}, Briggs (then responsible for the atomic bomb developement), got replaced by \emph{Ernest Lawrence}, a physicist from Berkeley. In December, after \emph{Pear Harbor} happenend, the project took of, now called \emph{The Manhattan Project}.

\subsection*{What was superior about the American bomb design?}
\emph{Eugene Wigner}, a refugee from Belgium, found the most efficient design for the shape of uranium inside a bomb: a sphere (Heisenberg used a flat sheet, which was the worst possible option). Wigner had an engineering degree, and Heisenberg did not, which provided many improvements to the atomic bomb.

\subsection*{What did the Allied forces do to win the race for the bomb?}
The USA had a pretty weak army compared to Germany (only 10th in the world). Germany also had the best engineers, which didn't make the race for the bomb any easier. The German effort would have to be sabotaged...

(Cliffhanger of this chapter. See next chapter, I guess..?)
