%%
% 15_creating-the-earth.tex
%
% @author: Patrick Scherler
%%
\newpage
\section{Creating the Earth}

\subsection*{1. What is the leading question of this chapter? (p. 185)}
How could E=mc2 operate to create the ordinary elements of our planet and daily life as well?

\subsection*{2. How did Hoyle explain the creation of the elements?}
If a star ever imploded, its center could reach close to 100 million degrees. That would be enough to squeeze even the larger nuclei of more massive elements together.

\subsection*{3. Where did he get his inspiration for the theory of implosion?}
He found out that the scientists of the Manhattan Project use implosion to get a full nuclear reaction of the plutonium when the bomb explodes. Implosion raises the pressure and temperature enough to do that.

\subsection*{4. What keeps our planet hot at the core and causes continental shifts (earthquakes)?}
Along with the initial heat left over from the impacts of Earth's creation, the radioactive blastings from the uranium and similar heavy elements have kept our planet's depths from cooling.

This high temperatures underneath the surface shift the thin continents on top to shape the surface of Earth.

\subsection*{5. What examples of modern human applications of E=mc2 are given?}
Atomic bombs were one of the first direct applications. Then nuclear submarines and power stations were built. E=mc2 continues at work in ordinary houses (smoke detectors, exit signs), hospitals (PET scans) and the archaeology (C-14 clock).